---
title: "210318_HeatmapsOfRelevantGenes_Pollak"
author: "Erin Osborne Nishimura"
date: "3/18/2021"
output: html_document
---

# Analysis: Heatmaps of key gene signature sets

The purpose of this analysis is to assess how gene signature sets are changing the Pollak et al dataset. 
Here is the Pollak et al dataset:

  * Pollak 

For this analysis, we used the following publications:

  * Creighton CJ. [A gene transcription signature of the Akt/mTOR pathway in clinical breast tumors](https://doi.org/10.1038/sj.onc.1210245). _Oncogene_ **2007**;26:4648–55.

  * Dry JR, Pavey S, Pratilas CA, Harbron C, Runswick S, Hodgson D, et al. [Transcriptional Pathway Signatures Predict MEK Addiction and Response to Selumetinib (AZD6244)](https://doi.org/10.1158/0008-5472.can-09-1577). _Cancer Res_ **2010**;70:2264–73.

  * Loboda A, Nebozhyn M, Klinghoffer R, Frazier J, Chastain M, Arthur W, et al. [A gene expression signature of RAS pathway dependence predicts response to PI3K and RAS pathway inhibitors and expands the population of RAS pathway activated tumors.](https://doi.org/10.1186/1755-8794-3-26) _Bmc Med Genomics_ **2010**;3:26.

  * Pratilas CA, Taylor BS, Ye Q, Viale A, Sander C, Solit DB, et al. [V600EBRAF is associated with disabled feedback inhibition of RAF–MEK signaling and elevated transcriptional output of the pathway.](https://doi.org/10.1073/pnas.0900780106) _Proc National Acad Sci_ **2009**;106:4519–24. 

  * Wagle M-C, Kirouac D, Klijn C, Liu B, Mahajan S, Junttila M, et al. [A transcriptional MAPK Pathway Activity Score (MPAS) is a clinically relevant biomarker in multiple cancer types](https://doi.org/10.1038/s41698-018-0051-4). _Npj Precis Oncol_ **2018**;2:7.

```{r, echo = FALSE}

# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install(version = "3.12")
# 
# library(BiocManager)
# BiocManager::install("ComplexHeatmap")
# I need to download X11 - XQuartz on Mac to use ComplexHeatmap


#if (!requireNamespace("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")
#BiocManager::install(version = "3.12")

#BiocManager::install("clusterProfiler")


```



## Step 1: Load packages
```{r load packges, warning = FALSE, message = FALSE}

library(RColorBrewer)
library(ggplot2)
library(dplyr)
library(pheatmap)
library(reshape2)
library(pryr)
```

## Step 2: Set Today's date:
```{r set today, echo = FALSE}
today <- format(Sys.Date(), "%y%m%d")
today
```

## Step 3: Import 
Import the lists of changing genes from the Pollak paper generated in the DESeq2 analysis.

```{r, echo = FALSE}

# Obtain the Mek, Ras, and Akt Files:
upIn827_1 <- read.table(file = "../01_input/210405_1_upIn827.txt", header = TRUE)
upIn1502_2 <- read.table(file = "../01_input/210405_2_upIn1502.txt", header = TRUE)
upIn131_3 <- read.table(file = "../01_input/210405_3_upIn131.txt", header = TRUE)
upInG166_4 <- read.table(file = "../01_input/210405_4_upInG166.txt", header = TRUE)
upInG179_5 <- read.table(file = "../01_input/210405_5_upInG179.txt", header = TRUE)

downIn827_6 <- read.table(file = "../01_input/210405_6_downIn827.txt", header = TRUE)
downIn1502_7 <- read.table(file = "../01_input/210405_7_downIn1502.txt", header = TRUE)
downIn131_8 <- read.table(file = "../01_input/210405_8_downIn131.txt", header = TRUE)
downInG166_9 <- read.table(file = "../01_input/210405_9_downInG166.txt", header = TRUE)
downInG179_10 <- read.table(file = "../01_input/210405_10_downInG179.txt", header = TRUE)

# import the list of all genes:
all_genes <- read.table(file = "../01_input/210215_unionAllGenes_.txt", header = TRUE)

dim(upIn827_1)
dim(upIn1502_2)
dim(upIn131_3)
dim(upInG166_4)
dim(upInG179_5)
```

## Step 4: Import the datasets from other published papers - 

```{r}
# Function - match and lookup
match_and_lookup_geneid <- function(published_genesetlist){
  all_genes[which(all_genes$external_gene_name %in% published_genesetlist),]$ensembl_gene_id
}

# Function - identify missing genes
identifyMissingGenes <- function(published_genesetlist){
  setdiff(published_genesetlist, all_genes$external_gene_name)
}

# Import Mek dataset
Mek18 <- read.table(file = "../01_input/MekEtAl18.txt", header = FALSE)
colnames(Mek18)[1] <- "genenames"
# get ensembl Gene IDs
mek18_ensbl_geneID <- match_and_lookup_geneid(Mek18$genenames)
identifyMissingGenes(Mek18$genenames)
missing_ensembl <- c("ENSG00000107104", "ENSG00000103994")
mek18_ensbl_geneID <- c(mek18_ensbl_geneID, missing_ensembl)
length(mek18_ensbl_geneID)
length(Mek18$genenames)


# Import Pratilas dataset
Pratilas_table <- read.table(file = "../01_input/Pratilis_genenames.txt", header = FALSE)
colnames(Pratilas_table)[1] <- "genenames"
# get ensembl Gene IDs
Pratilas_ensbl_geneID <- match_and_lookup_geneid(Pratilas_table$genenames)
identifyMissingGenes(Pratilas_table$genenames)
#missing
# "ALF"      "BXDC2"    "CD3EAP"   "FLJ10534" "HSPC111"  "IL8"      "KIR3DL2" 
# "LNK"      "PYCRL"
missing_ensmbl <- c("ENSG00000242441", "ENSG00000113460", "ENSG00000117877", "ENSG00000167721", "ENSG00000048162", "ENSG00000169429", "ENSG00000240403", "ENSG00000111252", "ENSG00000104524")
Pratilas_ensbl_geneID <- c(Pratilas_ensbl_geneID, missing_ensmbl)
length(Pratilas_ensbl_geneID)
length(Pratilas_table$genenames)

# Wagle et al -> 10 genes
vector_of_genes_wagle <- c("SPRY2", "SPRY4", "ETV4", "ETV5", "DUSP4", "DUSP6", "CCND1", "EPHA2", "EPHA4", "PHLDA1")
wagle_ensbl_geneID <- match_and_lookup_geneid(vector_of_genes_wagle)
identifyMissingGenes(vector_of_genes_wagle)
# no missing genes

# Import Creighton dataset
creighton_table <- read.table(file = "../01_input/Creighton_list.txt", header = FALSE)
colnames(creighton_table)[1] <- "genenames"
creighton_ensbl_geneID <- match_and_lookup_geneid(creighton_table$genenames)
identifyMissingGenes(creighton_table$genenames)
missing_ensmbl <- c("ENSG00000197451", "ENSG00000102225", "ENSG00000130202", "ENSG00000064601", "ENSG00000188229")
creighton_ensbl_geneID <- c(creighton_ensbl_geneID, missing_ensmbl)
length(creighton_ensbl_geneID)
length(creighton_table$genenames)


# Import Loboda dataset
lobodaUp_table <- read.table(file = "../01_input/Loboda_RASUp.txt", header = FALSE)
colnames(lobodaUp_table)[1] <- "genenames"
lobodaUp_ensbl_geneID <- match_and_lookup_geneid(lobodaUp_table$genenames)
identifyMissingGenes(lobodaUp_table$genenames)
# "C19orf10" "C20orf42" "EREG"     "IL8"      "NPAL1"    "PBEF1"    "PTGS2"    "RPRC1"    "SERPINB5" "TNS4"    
# "TUBA1"    "VEGF"  
missing_ensmbl <- c("ENSG00000074842","ENSG00000101311", "ENSG00000124882", "ENSG00000169429", "ENSG00000163293", "ENSG00000105835", "ENSG00000073756", "ENSG00000116871",  "ENSG00000206075", "ENSG00000131746", "ENSG00000127824", "ENSG00000112715")
lobodaUp_ensbl_geneID <- c(lobodaUp_ensbl_geneID, missing_ensmbl)
length(lobodaUp_ensbl_geneID)
length(lobodaUp_table$genenames)
```

Write a function called *make_contingency_table* that creates contingency tables for statistical analyses:

Write a function called *vector_of_p_values* that creates a vector of p-values generated from comparisons with the DESeq2 differentially expressed genes expressing the likeliness that a given gene signature set of genes are **enriched** amongst the list of genes.

Write a function called *vector_of_depletion_p_values* that creates a vector of p=values generated from comparisons with the DESeq2 differentially expressed genes expressing the likeliness that a given gene signature set of genes are **depleted** amongst the list of genes.

```{r, echo = FALSE}
make_contingency_table <- function(Published_list, RNAseqList, totalList) {
  intersectA <- length(intersect(Published_list, RNAseqList))
  NotFoundB <- length(Published_list) - intersectA
  MissedC <- length(RNAseqList) - intersectA
  NeitherD <- length(totalList) - sum(intersectA, NotFoundB, MissedC)
  
  table1 <- rbind(RNAseqChanging = c(intersectA, MissedC), RNAseqUnchanged = c(NotFoundB, NeitherD))
  colnames(table1) <- c("published", "Notpublished")
  return(as.table(table1))
}



# return a vector of P values assoicated with enrichment in each RNA-seq geneset
vector_of_p_values <- function(Published_list){
  p_value_vector <- c()
  table <- make_contingency_table(Published_list, upIn827_1$EnsembleGeneName, 
                                  all_genes$ensembl_gene_id)
  p_value_vector[1] <- fisher.test(table, alternative = "greater")$p.value

  table <- make_contingency_table(Published_list, upIn1502_2$EnsembleGeneName, 
                                  all_genes$ensembl_gene_id)
  p_value_vector[2] <- fisher.test(table, alternative = "greater")$p.value  
  
  table <- make_contingency_table(Published_list, upIn131_3$EnsembleGeneName, 
                                  all_genes$ensembl_gene_id)
  p_value_vector[3] <- fisher.test(table, alternative = "greater")$p.value   
  
  table <- make_contingency_table(Published_list, upInG166_4$EnsembleGeneName, 
                                  all_genes$ensembl_gene_id)
  p_value_vector[4] <- fisher.test(table, alternative = "greater")$p.value  
  
  table <- make_contingency_table(Published_list, upInG179_5$EnsembleGeneName, 
                                  all_genes$ensembl_gene_id)
  p_value_vector[5] <- fisher.test(table, alternative = "greater")$p.value   
  
  return(p_value_vector)
}


# return a vector of P values associated with depletion in each RNA-seq geneset
vector_of_depletion_p_values <- function(Published_list){
  p_value_vector <- c()
  table <- make_contingency_table(Published_list, downIn827_6$V1, all_genes$ensembl_gene_id)
  p_value_vector[1] <- fisher.test(table, alternative = "less")$p.value

  table <- make_contingency_table(Published_list, downIn1502_7$V1, all_genes$ensembl_gene_id)
  p_value_vector[2] <- fisher.test(table, alternative = "less")$p.value  
  
  table <- make_contingency_table(Published_list, downIn131_8$V1, all_genes$ensembl_gene_id)
  p_value_vector[3] <- fisher.test(table, alternative = "less")$p.value   

  table <- make_contingency_table(Published_list, downInG166_9$V1, all_genes$ensembl_gene_id)
  p_value_vector[4] <- fisher.test(table, alternative = "less")$p.value  
  
  table <- make_contingency_table(Published_list, downInG179_10$V1, all_genes$ensembl_gene_id)
  p_value_vector[5] <- fisher.test(table, alternative = "less")$p.value 
  return(p_value_vector)
}




```

Write a function called *calculate_percent_enrichment* that calculates the over-representation between a published gene signature and the gene sets from our RNA-seq analysis. Our gene sets are lists of genes that are Up compared to ARPE-19 or down compared to ARPE-19 at a 0.1 adjusted p-value threshold and a 0.25 minimum log fold change.


Write a function called *calculate_fold_enrichment* that calcualets the fold enrichment between a published genes signature set and the genesets from our RNA-seq analysis.

```{r, echo = FALSE}

calculate_percent_enrichment <- function(vector_of_published_genes){
  percent <- c()
  
  table <- make_contingency_table(vector_of_published_genes, upIn827_1$EnsembleGeneName, all_genes$ensembl_gene_id)
  percent[11] <- sum(table[,1])/sum(table)*100
  
  percent[1] <- sum(table[1,1])/sum(table[1,])*100
  
  table <- make_contingency_table(vector_of_published_genes, upIn1502_2$EnsembleGeneName, all_genes$ensembl_gene_id)
  percent[2] <- sum(table[1,1])/sum(table[1,])*100
  
  table <- make_contingency_table(vector_of_published_genes, upIn131_3$EnsembleGeneName, all_genes$ensembl_gene_id)
  percent[3] <- sum(table[1,1])/sum(table[1,])*100
  
  table <- make_contingency_table(vector_of_published_genes, upInG166_4$EnsembleGeneName, all_genes$ensembl_gene_id)
  percent[4] <- sum(table[1,1])/sum(table[1,])*100
  
  table <- make_contingency_table(vector_of_published_genes, upInG179_5$EnsembleGeneName, all_genes$ensembl_gene_id)
  percent[5] <- sum(table[1,1])/sum(table[1,])*100

  table <- make_contingency_table(vector_of_published_genes, downIn827_6$EnsembleGeneName, all_genes$ensembl_gene_id)
  percent[6] <- sum(table[1,1])/sum(table[1,])*100
  
  table <- make_contingency_table(vector_of_published_genes, downIn1502_7$EnsembleGeneName, all_genes$ensembl_gene_id)
  percent[7] <- sum(table[1,1])/sum(table[1,])*100
  
  table <- make_contingency_table(vector_of_published_genes, downIn131_8$EnsembleGeneName, all_genes$ensembl_gene_id)
  percent[8] <- sum(table[1,1])/sum(table[1,])*100
  
  table <- make_contingency_table(vector_of_published_genes, downInG166_9$EnsembleGeneName, all_genes$ensembl_gene_id)
  percent[9] <- sum(table[1,1])/sum(table[1,])*100
  
  table <- make_contingency_table(vector_of_published_genes, downInG179_10$EnsembleGeneName, all_genes$ensembl_gene_id)
  percent[10] <- sum(table[1,1])/sum(table[1,])*100

  names(percent) <- c("UpIn827", "UpIn1502", "UpIn131", "UpInG166", "UpInG179", "DownIn827", "DownIn1502", "DownIn131", "DownInG166", "DownInG179", "Genomewide")

  return(percent)
}

calculate_percent_enrichment(Pratilas_ensbl_geneID)

calculate_fold_enrichment <- function(vector_of_published_genes){
    percent <- c()
    foldEnrichment <- c()
    
    table <- make_contingency_table(vector_of_published_genes, upIn827_1$EnsembleGeneName, all_genes$ensembl_gene_id)
    percent[11] <- sum(table[,1])/sum(table)*100
    percent[1] <- sum(table[1,1])/sum(table[1,])*100
    foldEnrichment[1] <- percent[1] / percent[11]
    
    table <- make_contingency_table(vector_of_published_genes, upIn1502_2$EnsembleGeneName, all_genes$ensembl_gene_id)
    percent[2] <- sum(table[1,1])/sum(table[1,])*100
    foldEnrichment[2] <- percent[2] / percent[11]
    
    table <- make_contingency_table(vector_of_published_genes, upIn131_3$EnsembleGeneName, all_genes$ensembl_gene_id)
    percent[3] <- sum(table[1,1])/sum(table[1,])*100
    foldEnrichment[3] <- percent[3] / percent[11]
  
table <- make_contingency_table(vector_of_published_genes, upInG166_4$EnsembleGeneName, all_genes$ensembl_gene_id)
    percent[4] <- sum(table[1,1])/sum(table[1,])*100
    foldEnrichment[4] <- percent[4] / percent[11]
    
table <- make_contingency_table(vector_of_published_genes, upInG179_5$EnsembleGeneName, all_genes$ensembl_gene_id)
    percent[5] <- sum(table[1,1])/sum(table[1,])*100
    foldEnrichment[5] <- percent[5] / percent[11]

    table <- make_contingency_table(vector_of_published_genes, downIn827_6$EnsembleGeneName, all_genes$ensembl_gene_id)
    percent[6] <- sum(table[1,1])/sum(table[1,])*100
    foldEnrichment[6] <- percent[6] / percent[11]
    
    table <- make_contingency_table(vector_of_published_genes, downIn1502_7$EnsembleGeneName, all_genes$ensembl_gene_id)
    percent[7] <- sum(table[1,1])/sum(table[1,])*100
    foldEnrichment[7] <- percent[7] / percent[11]
    
    table <- make_contingency_table(vector_of_published_genes, downIn131_8$EnsembleGeneName, all_genes$ensembl_gene_id)
    percent[8] <- sum(table[1,1])/sum(table[1,])*100
    foldEnrichment[8] <- percent[8] / percent[11]
  
    table <- make_contingency_table(vector_of_published_genes, downInG166_9$EnsembleGeneName, all_genes$ensembl_gene_id)
    percent[9] <- sum(table[1,1])/sum(table[1,])*100
    foldEnrichment[9] <- percent[9] / percent[11]
    
    table <- make_contingency_table(vector_of_published_genes, downInG179_10$EnsembleGeneName, all_genes$ensembl_gene_id)
    percent[10] <- sum(table[1,1])/sum(table[1,])*100
    foldEnrichment[10] <- percent[10] / percent[11]
    names(foldEnrichment) <- c("UpIn827", "UpIn1502", "UpIn131", "UpInG166", "UpInG179", "DownIn827", "DownIn1502", "DownIn131", "DownInG166", "DownInG179")
  
    return(foldEnrichment)
}



calculate_fold_depletion <- function(vector_of_published_genes){
    percent <- c()
    foldDepletion <- c()
    
    table <- make_contingency_table(vector_of_published_genes, downIn827_6$EnsembleGeneName, all_genes$ensembl_gene_id)
    percent[11] <- sum(table[,1])/sum(table)*100
    percent[1] <- sum(table[1,1])/sum(table[1,])*100
    foldDepletion[1] <- percent[1] / percent[11]
    
    table <- make_contingency_table(vector_of_published_genes, downIn1502_7$EnsembleGeneName, all_genes$ensembl_gene_id)
    percent[2] <- sum(table[1,1])/sum(table[1,])*100
    foldDepletion[2] <- percent[2] / percent[11]
    
    table <- make_contingency_table(vector_of_published_genes, downIn131_8$EnsembleGeneName, all_genes$ensembl_gene_id)
    percent[3] <- sum(table[1,1])/sum(table[1,])*100
    foldDepletion[3] <- percent[3] / percent[11]
    
    table <- make_contingency_table(vector_of_published_genes, downInG166_9$EnsembleGeneName, all_genes$ensembl_gene_id)
    percent[4] <- sum(table[1,1])/sum(table[1,])*100
    foldDepletion[4] <- percent[4] / percent[11]
    
    table <- make_contingency_table(vector_of_published_genes, downInG179_10$EnsembleGeneName, all_genes$ensembl_gene_id)
    percent[5] <- sum(table[1,1])/sum(table[1,])*100
    foldDepletion[5] <- percent[5] / percent[11]

    table <- make_contingency_table(vector_of_published_genes, downIn827_6$EnsembleGeneName, all_genes$ensembl_gene_id)
    percent[6] <- sum(table[1,1])/sum(table[1,])*100
    foldDepletion[6] <- percent[6] / percent[11]
    
    table <- make_contingency_table(vector_of_published_genes, downIn1502_7$EnsembleGeneName, all_genes$ensembl_gene_id)
    percent[7] <- sum(table[1,1])/sum(table[1,])*100
    foldDepletion[7] <- percent[7] / percent[11]
    
    table <- make_contingency_table(vector_of_published_genes, downIn131_8$EnsembleGeneName, all_genes$ensembl_gene_id)
    percent[8] <- sum(table[1,1])/sum(table[1,])*100
    foldDepletion[8] <- percent[8] / percent[11]
    
    table <- make_contingency_table(vector_of_published_genes, downInG166_9$EnsembleGeneName, all_genes$ensembl_gene_id)
    percent[9] <- sum(table[1,1])/sum(table[1,])*100
    foldDepletion[9] <- percent[9] / percent[11]
    
    table <- make_contingency_table(vector_of_published_genes, downInG179_10$EnsembleGeneName, all_genes$ensembl_gene_id)
    percent[10] <- sum(table[1,1])/sum(table[1,])*100
    foldDepletion[10] <- percent[10] / percent[10]
  
    names(foldDepletion) <- c("UpIn827", "UpIn1502", "UpIn131", "UpInG166", "UpInG179", "DownIn827", "DownIn1502", "DownIn131", "DownInG166", "DownInG179")
  
    return(foldDepletion)
}

```

## Step 5: Gather statistics

```{r, echo = FALSE}
######################
# P-values
######################

# Generate an empty dataframe for the p-values calculating statistical significance of enrichment:
Pvalue_df <- data.frame(UpIn827 = numeric(),
                        UpIn1502 = numeric(),
                        UpIn131 = numeric(),
                        UpInG166 = numeric(),
                        UpInG179 = numeric(),
                        DownIn827 = numeric(),
                        DownIn1502 = numeric(),
                        DownIn131 = numeric(),
                        DownInG166 = numeric(),
                        DownInG179 = numeric())

# Save P-values of statistical enrichment to a dataframe called Pvalue_df
Pvalue_df[1,] <- vector_of_p_values(mek18_ensbl_geneID)
Pvalue_df[2,] <- vector_of_p_values(Pratilas_ensbl_geneID)
Pvalue_df[3,] <- vector_of_p_values(wagle_ensbl_geneID)
Pvalue_df[4,] <- vector_of_p_values(creighton_ensbl_geneID)
Pvalue_df[5,] <- vector_of_p_values(lobodaUp_ensbl_geneID)

rownames(Pvalue_df)[1] <- "Dry_et_al_mek_18"
rownames(Pvalue_df)[2] <- "Pratilas_et_al_mek_50"
rownames(Pvalue_df)[3] <- "Wagle_et_al_mek_10"
rownames(Pvalue_df)[4] <- "Creighton_et_al_akt_57"
rownames(Pvalue_df)[5] <- "Loboda_ras_101"

Pvalue_df

######################
# Percent
######################

# Generate an empty dataframe for the percentages:
percent_df <- data.frame(UpIn827 = numeric(),
                        UpIn1502 = numeric(),
                        UpIn131 = numeric(),
                        UpInG166 = numeric(),
                        UpInG179 = numeric(),
                        DownIn827 = numeric(),
                        DownIn1502 = numeric(),
                        DownIn131 = numeric(),
                        DownInG166 = numeric(),
                        DownInG179 = numeric(),
                        Genomewide = numeric())

# Save percents to a dataframe called percent_df
percent_df[1,] <- calculate_percent_enrichment(mek18_ensbl_geneID)
percent_df[2,] <- calculate_percent_enrichment(Pratilas_ensbl_geneID)
percent_df[3,] <- calculate_percent_enrichment(wagle_ensbl_geneID)
percent_df[4,] <- calculate_percent_enrichment(creighton_ensbl_geneID)
percent_df[5,] <- calculate_percent_enrichment(lobodaUp_ensbl_geneID)

rownames(percent_df)[1] <- "Dry_et_al_mek_18"
rownames(percent_df)[2] <- "Pratilas_et_al_mek_50"
rownames(percent_df)[3] <- "Wagle_et_al_mek_10"
rownames(percent_df)[4] <- "Creighton_et_al_akt_57"
rownames(percent_df)[5] <- "Loboda_ras_101"

percent_df

######################
# Fold Enrichment
######################

# Generate an empty dataframe for the fold enrichment:
df_enrich <- data.frame(UpIn827 = numeric(),
                        UpIn1502 = numeric(),
                        UpIn131 = numeric(),
                        UpInG166 = numeric(),
                        UpInG179 = numeric(),
                        DownIn827 = numeric(),
                        DownIn1502 = numeric(),
                        DownIn131 = numeric(),
                        DownInG166 = numeric(),
                        DownInG179 = numeric())

# Gather statistics: Fold enrichment - enrichment each list
df_enrich[1,] <- calculate_fold_enrichment(mek18_ensbl_geneID)
df_enrich[2,] <- calculate_fold_enrichment(Pratilas_ensbl_geneID)
df_enrich[3,] <- calculate_fold_enrichment(wagle_ensbl_geneID)
df_enrich[4,] <- calculate_fold_enrichment(creighton_ensbl_geneID)
df_enrich[5,] <- calculate_fold_enrichment(lobodaUp_ensbl_geneID)

rownames(df_enrich)[1] <- "Dry_et_al_mek_18"
rownames(df_enrich)[2] <- "Pratilas_et_al_mek_50"
rownames(df_enrich)[3] <- "Wagle_et_al_mek_10"
rownames(df_enrich)[4] <- "Creighton_et_al_akt_57"
rownames(df_enrich)[5] <- "Loboda_ras_101"

df_enrich

######################
# Fold Depletion
######################

# Generate an empty dataframe for the fold depletion:
df_deplete <- data.frame(UpIn827 = numeric(),
                        UpIn1502 = numeric(),
                        UpIn131 = numeric(),
                        UpInG166 = numeric(),
                        UpInG179 = numeric(),
                        DownIn827 = numeric(),
                        DownIn1502 = numeric(),
                        DownIn131 = numeric(),
                        DownInG166 = numeric(),
                        DownInG179 = numeric())

# Gather statistics: Fold enrichment - enrichment each list
df_deplete[1,] <- calculate_fold_depletion(mek18_ensbl_geneID)
df_deplete[2,] <- calculate_fold_depletion(Pratilas_ensbl_geneID)
df_deplete[3,] <- calculate_fold_depletion(wagle_ensbl_geneID)
df_deplete[4,] <- calculate_fold_depletion(creighton_ensbl_geneID)
df_deplete[5,] <- calculate_fold_depletion(lobodaUp_ensbl_geneID)

rownames(df_deplete)[1] <- "Dry_et_al_mek_18"
rownames(df_deplete)[2] <- "Pratilas_et_al_mek_50"
rownames(df_deplete)[3] <- "Wagle_et_al_mek_10"
rownames(df_deplete)[4] <- "Creighton_et_al_akt_57"
rownames(df_deplete)[5] <- "Loboda_ras_101"

df_deplete

```

## Step 6: Barplots of enrichment/Depletion

```{r}

# Plot barplot of Dry et al, Mek
p_mek18 %<a-% {
  barplot(rev(as.matrix(percent_df[1,1:5])), horiz = TRUE, las = 2,
          xlim = c(0,4),
          main = "Dry et al., 2010; 18 MEK genes",
          xlab = "percentage",
          names.arg = rev(colnames(percent_df[1,1:5])) )
  abline(v=as.numeric(percent_df[1,11]), col="red",lty=2)
}

p_mek18

# Plot of Pratilas et al., Mek
p_pratilas %<a-% {
  barplot(rev(as.matrix(percent_df[2,1:5])), horiz = TRUE, las = 2,
          xlim = c(0,4),
          main = "Pratilas, MEK genes",
          xlab = "percentage",
          names.arg = rev(colnames(percent_df[2,1:5])) )
  abline(v=as.numeric(percent_df[1,11]), col="red",lty=2)
}

p_pratilas

# Plot of Wagle et al., Mek
p_wagle %<a-% {
  barplot(rev(as.matrix(percent_df[3,1:5])), horiz = TRUE, las = 2,
          xlim = c(0,4),
          main = "Wagle, MEK genes",
          xlab = "percentage",
          names.arg = rev(colnames(percent_df[3,1:5])) )
  abline(v=as.numeric(percent_df[1,11]), col="red",lty=2)
}

p_wagle

# Plot of Wagle et al., Mek
p_creighton %<a-% {
  barplot(rev(as.matrix(percent_df[4,1:5])), horiz = TRUE, las = 2,
          xlim = c(0,4),
          main = "Creighton et al, AKT genes",
          xlab = "percentage",
          names.arg = rev(colnames(percent_df[4,1:5])) )
  abline(v=as.numeric(percent_df[1,11]), col="red",lty=2)
}

p_creighton


p_loboda %<a-% {
  barplot(rev(as.matrix(percent_df[5,1:5])), horiz = TRUE, las = 2,
          xlim = c(0,4),
          main = "Loboda et al, RAS genes",
          xlab = "percentage",
          names.arg = rev(colnames(percent_df[5,1:5])) )
  abline(v=as.numeric(percent_df[1,11]), col="red",lty=2)
}

p_loboda

filename = paste("../03_output/", today, "_Dry18_barplot.pdf", sep = "")
pdf(file = filename, width = 5, height = 4)
  p_mek18
  dev.off()

```

## Step 7: Dot plot:

```{r}


# Subset the enrichment p-values to those I will use in a figure:
matrix_of_Pvalues <- -log(Pvalue_df[c(4,2,1,3,5),1:5])

# Gather the fold change data 
df_enrich_reorder <- df_enrich[c(4,2,1,3,5),1:5]
melted_foldEnrich <- melt(df_enrich_reorder, value.name = "fold_enrich")

# Reshape the enrichment p-values into a long format dataset:
long_Pvalues <- as.matrix(matrix_of_Pvalues) %>% 
  melt() %>%
  mutate(RNAseq_num = c(rep(1, 5), rep(2,5), rep(3,5), rep(4,5), rep(5,5))) %>%
  mutate(PubList = c(rep(c(1,2,3,4,5), 5))) %>%
  mutate(foldenrich = melted_foldEnrich$fold_enrich)

long_Pvalues

## The 0 values in the depletion matrix cause some problems numerically. Change these to a very low number.
#depletion_matrix
#depletion_matrix[which(depletion_matrix$folddeplete == 0),6] <- 0.00000000000000000001
        
# Plot the data - enrichment
plot1 <- ggplot(long_Pvalues, aes(x=RNAseq_num, y=PubList, color=log(foldenrich))) + 
  geom_point(aes(size=value)) +
  scale_colour_gradient2(midpoint = 0, low = "navy", mid = "grey", high = "magenta", limits = c(-3, 3)) +
  xlim(0.5,5.5) +
  ylim(5.5, 0.5) +
  scale_size(range = c(1, 20)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), panel.grid.major = )

plot1
help("scale_colour_gradient2")
long_Pvalues

filename = paste("../03_output/", today, "_pvalue_enrich_dotplot.pdf", sep = "")
pdf(file = filename, width = 6, height = 4.5)
plot1
dev.off()


Pvalue_df
```










########################

## Step 5: Select genes in different branches of the RAS/AKT/MEK pathways:

These are the genes that were altered as part of the transformation process.

```{r, echo = FALSE}
top_genes <- c("CCND1", "CDK4", "TERT", "TP53", "AKT1", "MAP2K1", "HRAS") 
top_genes2 <- c("CCND1_transgene", "CDK4r24c_transgene", "HRASV12_transgene",
               "MEKDD_transgene", "murine_p53DD_transgene", "MYR_FLAG_AKT_transgene", 
               "TERT_transgene")

# subset dataset
topgenes_df <- subset_counts_by_external_df(rlog_counts, top_genes)
topgenes_df2 <- subset_counts_by_ensembl_df(rlog_counts, top_genes2)
topgenes_df2$external_gene_name <- topgenes_df2$EnsembleGeneName

topgenes_df <- rbind(topgenes_df, topgenes_df2)

# make a matrix
topgenes_matrix <- return_labeled_matrix(topgenes_df)
topgenes_selection_matrix <- topgenes_matrix[c(1,1,12,5,4,11,7),]

# show heatmap
pheatmap(topgenes_selection_matrix,
         scale = "row",
         cluster_cols = FALSE,
         cluster_rows = FALSE)

# save heatmap
filename = paste("../03_output/", today, "_3_Top_heatmap.pdf", sep = "")
pdf(filename, width = 5, height = 3)
pheatmap(topgenes_selection_matrix,
         scale = "row",
         cluster_cols = FALSE,
         cluster_rows = FALSE)
dev.off()
dev.off()

```


## Step 5: Compare to Dry et al., 2010 - Mek-18 gene signature (18 genes)

This a gene signature set of genes that are predicted to be up-regulated with Mek-pathway activation. 

```{r, echo = FALSE}
Mek18 <- read.table(file = "../01_input/MekEtAl18.txt", header = FALSE)
colnames(Mek18)[1] <- "genenames"

# Subset the data
Mek_18_df <- subset_counts_by_external_df(rlog_counts, Mek18$genenames)

# Find missing genes
missing <- identify_missing_genes(Mek_18_df, Mek18$genenames)
#missing
  # "ANKRD15" "ZFP106"
missing_ensembl <- c("ENSG00000107104", "ENSG00000103994")

# Attach missing genes
Mek_18_df <- rbind(Mek_18_df, subset_counts_by_ensembl_df(rlog_counts, missing_ensembl))

# Dry et al., --> 18 genes
Mek18_Matrix <- return_labeled_matrix(Mek_18_df)

# Calculate enrichment
percent_enrich_Mek18 <- calculate_percent_enrichment(Mek_18_df$EnsembleGeneName)
fold_enrich_Mek18 <- calculate_fold_enrichment(Mek_18_df$EnsembleGeneName)
fold_enrich_Mek18

fold_deplete_Mek18 <- calculate_fold_depletion(Mek_18_df$EnsembleGeneName)
fold_deplete_Mek18

# Generate a barplot 
p_mek18 %<a-% {
  barplot(rev(percent_enrich_Mek18), horiz = TRUE, las = 2,
          xlim = c(0,4),
          main = "Dry et al., 2010; 18 MEK genes",
          xlab = "percentage") 
  abline(v=as.numeric(percent_enrich_Mek18[7]), col="red",lty=2)
}

p_mek18

filename = paste("../03_output/", today, "_Dry18_barplot.pdf", sep = "")
pdf(file = filename, width = 5, height = 4)
  p_mek18
  dev.off()
```

Heatmap of genes in the Mek 18 gene list: 

```{r, echo = FALSE}
pheatmap(Mek18_Matrix,
         scale = "row", 
         cluster_cols = FALSE)


filename = paste("../03_output/", today, "_1_Mek18_heatmap.pdf", sep = "")
pdf(filename, width = 5, height = 5)
pheatmap(Mek18_Matrix,
         scale = "row", 
         cluster_cols = FALSE)
dev.off()

```

## Step 6: Pratilas et al. 2009, Mek-52 gene signature (52 genes)

Plot the heatmap from Pratilas et al., 52 genes

These are predicted to be up-regulated with Mek-pathway activation. 

```{r heatmap, echo = FALSE}

# Import Pratilas dataset
Pratilas_table <- read.table(file = "../01_input/Pratilis_genenames.txt", header = FALSE)
colnames(Pratilas_table)[1] <- "genenames"

# subset dataset
pratilas_df <- subset_counts_by_external_df(rlog_counts, Pratilas_table$genenames)

# identify missing genes
missing <- identify_missing_genes(pratilas_df, Pratilas_table$genenames)
#missing
# "ALF"      "BXDC2"    "CD3EAP"   "FLJ10534" "HSPC111"  "IL8"      "KIR3DL2" 
# "LNK"      "PYCRL"
missing_ensmbl <- c("ENSG00000242441", "ENSG00000113460", "ENSG00000117877", "ENSG00000167721", "ENSG00000048162", "ENSG00000169429", "ENSG00000240403", "ENSG00000111252", "ENSG00000104524")

# Attach the missing genes
pratilas_df <- rbind(pratilas_df, subset_counts_by_ensembl_df(rlog_counts, missing_ensmbl))

# make a matrix
pratilas_matrix <- return_labeled_matrix(pratilas_df)

#Gather enrichments & depletions
percent_enrich_pratilas <- calculate_percent_enrichment(pratilas_df$EnsembleGeneName)
fold_enrich_pratilas <- calculate_fold_enrichment(pratilas_df$EnsembleGeneName)
fold_enrich_pratilas
fold_deplete_pratilas <- calculate_fold_depletion(pratilas_df$EnsembleGeneName)
fold_deplete_pratilas

# Print pratilas plot in output file
barplot(rev(percent_enrich_pratilas), horiz = TRUE, las = 2, xlim = c(0,4), main = "Pratilas - 50 - genes")
abline(v=as.numeric(percent_enrich_pratilas[7]), col="red",lty=2)

# print Pratilas plot as .pdf
filename = paste("../03_output/", today, "_Pratilas_barplot.pdf", sep = "")
pdf(file = filename, width = 5, height = 4)
barplot(rev(percent_enrich_pratilas), horiz = TRUE, las = 2, xlim = c(0,4), main = "Pratilas - 50 - genes")
abline(v=as.numeric(percent_enrich_pratilas[7]), col="red",lty=2)
dev.off()

```

Plot the heatmap of genes in the Pratilas gene set: 
```{r, echo = FALSE}
pheatmap(pratilas_matrix,
         scale = "row",
         cluster_cols = FALSE)

filename = paste("../03_output/", today, "_2_Pratilas_Mek_heatmap.pdf", sep = "")
pdf(filename, width = 5, height = 7)
pheatmap(pratilas_matrix,
         scale = "row",
         cluster_cols = FALSE)
dev.off()

```




## Step 7: Wagle et al., 2018. MEK pathway (10 genes)

These are predicted to be up-regulated with Mek-pathway activation. 

```{r, echo = FALSE}

# Wagle et al -> 10 genes
vector_of_genes_wagle <- c("SPRY2", "SPRY4", "ETV4", "ETV5", "DUSP4", "DUSP6", "CCND1", "EPHA2", "EPHA4", "PHLDA1")

# subset dataset
wagle_df <- subset_counts_by_external_df(rlog_counts, vector_of_genes_wagle)

# identify missing genes
missing <- identify_missing_genes(wagle_df, vector_of_genes_wagle)
# no missing genes

# Gather enrichment & depletion rates
percent_enrich_wagle <- calculate_percent_enrichment(wagle_df$EnsembleGeneName)
fold_enrich_wagle <- calculate_fold_enrichment(wagle_df$EnsembleGeneName)
fold_enrich_wagle
fold_deplete_wagle <- calculate_fold_depletion(wagle_df$EnsembleGeneName)
fold_deplete_wagle

# plot 
barplot(rev(percent_enrich_wagle), horiz = TRUE, las = 2, xlim = c(0,4), main = "Wagle et al., - 10 Mek genes")
abline(v=as.numeric(percent_enrich_wagle[7]), col="red",lty=2)

# Save the plot
filename = paste("../03_output/", today, "_Wagle_barplot.pdf", sep = "")
pdf(file = filename, width = 5, height = 4)
barplot(rev(percent_enrich_wagle), horiz = TRUE, las = 2, xlim = c(0,4), main = "Wagle et al., - 10 Mek genes")
abline(v=as.numeric(percent_enrich_wagle[7]), col="red",lty=2)
dev.off()

```

Plot the heatmap of the genes in the Wagle geneset:

```{r, echo = FALSE}

# Make a matrix
wagle_matrix <- return_labeled_matrix(wagle_df)

# create a heatmap
pheatmap(wagle_matrix,
         scale = "row",
         cluster_cols = FALSE)

# Save the heatmap
filename = paste("../03_output/", today, "_4_WagleMek_heatmap.pdf", sep = "")
pdf(filename, width = 5, height = 3.5)
pheatmap(wagle_matrix,
         scale = "row", 
         cluster_cols = FALSE)
dev.off()

```

## Step 8: Creighton et al, 2007. AKT/mTOR pathway (57)

These are predicted to be up-regulated with AKT activation specifically in breast tumors 

```{r, echo = FALSE}
creighton_table <- read.table(file = "../01_input/Creighton_list.txt", header = FALSE)

# subset dataset
creighton_df <- subset_counts_by_external_df(rlog_counts, creighton_table$V1)

# identify missing genes
missing <- identify_missing_genes(creighton_df, creighton_table$V1)
#missing
# [1] "HNRPAB" "PCTK1"  "PVRL2"  "PPGB"   "TUBB2C"
missing_ensmbl <- c("ENSG00000197451", "ENSG00000102225", "ENSG00000130202", "ENSG00000064601", "ENSG00000188229")

# Attach missing genes
# Attach the missing genes
creighton_df <- rbind(creighton_df, subset_counts_by_ensembl_df(rlog_counts, missing_ensmbl))

# make a matrix
creighton_matrix <- return_labeled_matrix(creighton_df)

# Gather enrichment & depletion:
percent_enrich_creighton <- calculate_percent_enrichment(creighton_df$EnsembleGeneName)
fold_enrich_creighton <- calculate_fold_enrichment(creighton_df$EnsembleGeneName)
fold_enrich_creighton
fold_deplete_creighton <- calculate_fold_depletion(creighton_df$EnsembleGeneName)
fold_deplete_creighton

# barplot
barplot(rev(percent_enrich_creighton), horiz = TRUE, las = 2, xlim = c(0,4), main = "Creighton - 57 - Akt")
abline(v=as.numeric(percent_enrich_creighton[7]), col="red",lty=2)

# save barplot
filename = paste("../03_output/", today, "_Creighton_barplot.pdf", sep = "")
pdf(file = filename, width = 5, height = 4)
barplot(rev(percent_enrich_creighton), horiz = TRUE, las = 2, xlim = c(0,4), main = "Creighton - 57 - Akt")
abline(v=as.numeric(percent_enrich_creighton[7]), col="red",lty=2)
dev.off()

```

Plot the heatmap of genes in the Creighton gene set (AKT pathway):

```{r, echo = FALSE}
pheatmap(creighton_matrix,
         scale = "row",
         cluster_cols = FALSE)

filename = paste("../03_output/", today, "_5_CreightonAkt_heatmap.pdf", sep = "")
pdf(filename, width = 5, height = 7)
pheatmap(creighton_matrix,
         scale = "row", 
         cluster_cols = FALSE)
dev.off()

```

## Step 9: Loboda et al., 2010. RAS pathway - activated

These are predicted to be up-regulated with RAS activation. 105 genes.

```{r, echo = FALSE}
lobodaUp_table <- read.table(file = "../01_input/Loboda_RASUp.txt", header = FALSE)

# subset dataset
lobodaUp_df <- subset_counts_by_external_df(rlog_counts, lobodaUp_table$V1)

# identify missing genes
missing <- identify_missing_genes(lobodaUp_df, lobodaUp_table$V1)
#missing
#[1] "C19orf10" "C20orf42" "EREG"     "IL8"      "NPAL1"    "PBEF1"    "PTGS2"   
#[8] "RPRC1"    "SERPINB5" "TNS4"     "TUBA1"    "VEGF"  
missing_ensmbl <- c("ENSG00000074842","ENSG00000101311", "ENSG00000124882", "ENSG00000169429", "ENSG00000163293", "ENSG00000105835", "ENSG00000073756", "ENSG00000116871",  "ENSG00000206075", "ENSG00000131746", "ENSG00000127824", "ENSG00000112715")

# Attach missing genes
# Attach the missing genes
lobodaUp_df <- rbind(lobodaUp_df, subset_counts_by_ensembl_df(rlog_counts, missing_ensmbl))

# make a matrix
lobodaUp_matrix <- return_labeled_matrix(lobodaUp_df)

# Gather statistics: fold enrichment and depletion
percent_enrich_loboda <- calculate_percent_enrichment(lobodaUp_df$EnsembleGeneName)
fold_enrich_loboda <- calculate_fold_enrichment(lobodaUp_df$EnsembleGeneName)
fold_enrich_loboda
fold_deplete_loboda <- calculate_fold_depletion(lobodaUp_df$EnsembleGeneName)
fold_deplete_loboda

# Plot
barplot(rev(percent_enrich_loboda), horiz = TRUE, las = 2, main = "Loboda - 101 - Ras")
abline(v=as.numeric(percent_enrich_loboda[7]), col="red",lty=2)

# save barplot
filename = paste("../03_output/", today, "_Loboda_barplot.pdf", sep = "")
pdf(file = filename, width = 5, height = 4)
barplot(rev(percent_enrich_loboda), horiz = TRUE, las = 2, main = "Loboda - 101 - Ras")
abline(v=as.numeric(percent_enrich_loboda[7]), col="red",lty=2)
dev.off()

```

Plot the heatmap of genes in the Loboda Gene signature (RAS pathway):

```{r, echo = FALSE}
# plot the heatmap
pheatmap(as.matrix(lobodaUp_matrix),
         scale = "row",
         cluster_cols = FALSE)

filename = paste("../03_output/", today, "_6_Loboda_heatmap.pdf", sep = "")
pdf(filename, width = 5, height = 7)
pheatmap(as.matrix(lobodaUp_matrix),
         scale = "row",
         cluster_cols = FALSE)
dev.off()

```

## Step 10: Gather up the P-values for enrichment and depletion rates

```{r, echo = FALSE}
# Generate an empty dataframe for the enrichment p-values:
Pvalue_df <- data.frame(UpInAkt = numeric(),
                        UpInMek = numeric(),
                        UpinRas = numeric(),
                        DowninAkt = numeric(),
                        DowninMek = numeric(),
                        DowninRas = numeric())

# Save Dry et al., Mek18 data to the dataframe:
Pvalue_df[1,] <- vector_of_p_values(Mek_18_df$EnsembleGeneName)
rownames(Pvalue_df)[1] <- "Dry_et_al_mek_18"

# Generate an empty dataframe for the depletion p-values:
Pvalue_df_deplete <- data.frame(UpInAkt = numeric(),
                        UpInMek = numeric(),
                        UpinRas = numeric(),
                        DowninAkt = numeric(),
                        DowninMek = numeric(),
                        DowninRas = numeric())

Pvalue_df_deplete[1,] <- vector_of_depletion_p_values(Mek_18_df$EnsembleGeneName)
rownames(Pvalue_df_deplete)[1] <- "Dry_et_al_mek_18"



#Gather statistics
Pvalue_df[2,] <- vector_of_p_values(pratilas_df$EnsembleGeneName)
rownames(Pvalue_df)[2] <- "Pratilas_et_al_mek_50"
Pvalue_df_deplete[2,] <- vector_of_depletion_p_values(pratilas_df$EnsembleGeneName)
rownames(Pvalue_df_deplete)[2] <- "Pratilas_et_al_mek_50"

#Gather statistics
Pvalue_df[3,] <- vector_of_p_values(wagle_df$EnsembleGeneName)
rownames(Pvalue_df)[3] <- "Wagle_et_al_mek_10"
Pvalue_df_deplete[3,] <- vector_of_depletion_p_values(wagle_df$EnsembleGeneName)
rownames(Pvalue_df_deplete)[3] <- "Wagle_et_al_mek_10"

#Gather statistics
Pvalue_df[4,] <- vector_of_p_values(creighton_df$EnsembleGeneName)
rownames(Pvalue_df)[4] <- "Creighton_et_al_akt_57"
Pvalue_df_deplete[4,] <- vector_of_depletion_p_values(creighton_df$EnsembleGeneName)
rownames(Pvalue_df_deplete)[4] <- "Creighton_et_al_akt_57"

#Gather statistics
Pvalue_df[5,] <- vector_of_p_values(lobodaUp_df$EnsembleGeneName)
rownames(Pvalue_df)[5] <- "Loboda_ras_101"
Pvalue_df_deplete[5,] <- vector_of_depletion_p_values(lobodaUp_df$EnsembleGeneName)
rownames(Pvalue_df_deplete)[5] <- "Loboda_ras_101"

```

Here are the p-values reflecting the chance that there is enrichment in a gene signature within our geneset lits:

```{r, echo = FALSE}
Pvalue_df

```

Here are the p-values reflecting the chance that there is depletion in a gene signature within our geneset lits:

```{r, echo = FALSE}
Pvalue_df_deplete

```


## Step 11: Create dotplots for enrichment - fold change and p-values and create dotplots for depletion - fold chang and p-values


```{r, echo = FALSE}

# Subset the enrichment p-values to those I will use in a figure:
matrix_of_Pvalues <- -log(Pvalue_df)[c(4,2,1,3,5),1:3]
matrix_of_Pvalues
# Subset the depletion p-values to those I will use in a figure:
matrix_of_depletion_Pvalues <- -log(Pvalue_df_deplete)[c(4,2,1,3,5),4:6]

# Gather the fold change data 
allFoldChange <- c(
  fold_enrich_creighton[1],
  fold_enrich_pratilas[1],
  fold_enrich_Mek18[1],
  fold_enrich_wagle[1],
  fold_enrich_loboda[1],
    fold_enrich_creighton[2],
  fold_enrich_pratilas[2],
  fold_enrich_Mek18[2],
  fold_enrich_wagle[2],
  fold_enrich_loboda[2],
      fold_enrich_creighton[3],
  fold_enrich_pratilas[3],
  fold_enrich_Mek18[3],
  fold_enrich_wagle[3],
  fold_enrich_loboda[3])

# Gather the depletion data
allFoldDepletion <- c(
  fold_deplete_creighton[1],
  fold_deplete_pratilas[1],
  fold_deplete_Mek18[1],
  fold_deplete_wagle[1],
  fold_deplete_loboda[1],
    fold_deplete_creighton[2],
  fold_deplete_pratilas[2],
  fold_deplete_Mek18[2],
  fold_deplete_wagle[2],
  fold_deplete_loboda[2],
      fold_deplete_creighton[3],
  fold_deplete_pratilas[3],
  fold_deplete_Mek18[3],
  fold_deplete_wagle[3],
  fold_deplete_loboda[3])

# Reshape the enrichment p-values into a long format dataset:
long_Pvalues <- as.matrix(matrix_of_Pvalues) %>% 
  melt() %>%
  mutate(RNAseq_num = c(rep(1, 5), rep(2,5), rep(3,5))) %>%
  mutate(PubList = c(rep(c(1,2,3,4,5), 3))) %>%
  mutate(foldenrich = allFoldChange)

# Reshape the depletion p-values into a long format dataset:
depletion_matrix <- as.matrix(matrix_of_depletion_Pvalues) %>%
  melt() %>%
  mutate(RNAseq_num = c(rep(1, 5), rep(2,5), rep(3,5))) %>%
  mutate(PubList = c(rep(c(1,2,3,4,5), 3))) %>%
  mutate(folddeplete = allFoldDepletion)

# The 0 values in the depletion matrix cause some problems numerically. Change these to a very low number.
depletion_matrix
depletion_matrix[which(depletion_matrix$folddeplete == 0),6] <- 0.00000000000000000001
        
# Plot the data - enrichment
plot1 <- ggplot(long_Pvalues, aes(x=RNAseq_num, y=PubList, color=log(foldenrich))) + 
  geom_point(aes(size=value)) +
  scale_colour_gradient2(midpoint = 1, low = "navy", mid = "magenta4", high = "hotpink") +
  xlim(0.5,3.5) +
  ylim(5.5, 0.5) +
  scale_size(range = c(1, 15)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), panel.grid.major = )

plot1

# Plot the data - depletion
plot2 <- ggplot(depletion_matrix, aes(x=RNAseq_num, y=PubList, color=log(folddeplete))) + 
  geom_point(aes(size=value)) +
  scale_colour_gradient2(midpoint = 1, low = "cyan", mid = "blue", high = "navy") +
  xlim(0.5,3.5) +
  ylim(5.5, 0.5) +
  scale_size(range = c(1, 3)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), panel.grid.major = )

plot2

filename = paste("../03_output/", today, "_pvalue_enrich_dotplot.pdf", sep = "")
pdf(file = filename, width = 6, height = 4.5)
plot1
dev.off()


filename = paste("../03_output/", today, "_pvalue_deplete_dotplot.pdf", sep = "")
pdf(file = filename, width = 6, height = 4.5)
plot2
dev.off()

```



Session info:

```{r, echo = TRUE}
sessionInfo()
citation()
```

